---
title: "Extracting Common ADSL Variables Across Domains"
author: "Clinical Data Science Team"
vignette: >
  %\VignetteIndexEntry{Extract Common ADSL Variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
``` 

# Introduction

In clinical data analysis, the ADSL (Subject-Level Analysis Dataset) often contains critical variables that are reused across various analysis domains (e.g., ADAE, ADLB, ADLBC). The `extract_common_adsl_variables()` function in the **adrgOS** package helps identify and summarize these common variables, providing insights into variable provenance and consistency across domains.

# Installation

You can install **adrgOS** from GitHub:

```r
# install.packages("remotes")
remotes::install_github("phuse-org/adrgOS")
```  

# Usage

## Extracting variable information from define.xml

First, extract variable metadata from your Define-XML file using `extract_variable_info_from_define()`. If you have a Define-XML file in your package (`inst/define.xml`), you can locate it with `system.file()`.

```{r define-extract}
library(adrgOS)
define_path <- system.file("define.xml", package = "adrgOS")
if (file.exists(define_path)) {
  variable_info <- extract_variable_info_from_define(define_path)
} else {
  stop("Define-XML file not found in the package.")
}
```

## Extract common ADSL variables

Use `extract_common_adsl_variables()` to identify variables present in ADSL and other domains.

```{r extract-common-vars}
common_vars <- extract_common_adsl_variables(
  variable_info_df = variable_info,
  adsl_dataset_name = "ADSL",      # Name of the ADSL dataset
  include_sas_info    = TRUE,        # Include SAS field and length info
  min_domains_count   = 1,           # Minimum number of domains
  sort_by             = "domains_count"  # Sort by number of domains
)

# View the first few rows
head(common_vars)
```

# Customizing the results

### Filtering by domain frequency

To restrict to variables that appear in at least 3 other domains:

```{r filter-min-domains}
high_freq_vars <- extract_common_adsl_variables(
  variable_info_df = variable_info,
  min_domains_count = 3
)
head(high_freq_vars)
```

### Excluding SAS-specific information

If you prefer a simpler output without SAS field names and lengths, set `include_sas_info = FALSE`:

```{r exclude-sas-info}
vars_no_sas <- extract_common_adsl_variables(
  variable_info_df = variable_info,
  include_sas_info = FALSE,
  sort_by = "variable_name"
)
head(vars_no_sas)
```

# Working with sample data

When testing or demonstrating functionality, you can create a simple sample data frame:

```{r sample-data}
sample_data <- data.frame(
  Name = c("STUDYID", "USUBJID", "AGE", "SEX", 
           "STUDYID", "USUBJID", "AVAL", 
           "STUDYID", "USUBJID", "PARAMCD"),
  Dataset = c("ADSL", "ADSL", "ADSL", "ADSL", 
              "ADAE", "ADAE", "ADAE", 
              "ADLBC", "ADLBC", "ADLBC"),
  DataType = c("text", "text", "integer", "text",
               "text", "text", "float",
               "text", "text", "text"),
  Length = c("12", "20", "3", "1",
             "12", "20", "8",
             "12", "20", "8"),
  Description = c("Study Identifier", "Unique Subject ID", "Age", "Sex",
                  "Study Identifier", "Unique Subject ID", "Analysis Value",
                  "Study Identifier", "Unique Subject ID", "Parameter Code"),
  Origin = c("Assigned", "Assigned", "Collected", "Collected",
             "Assigned", "Assigned", "Derived", "Assigned", "Assigned", "Derived"),
  stringsAsFactors = FALSE
)

# Extract from sample data
result <- extract_common_adsl_variables(sample_data)
print(result)
```

# Conclusion

The **adrgOS** vignette demonstrated how to use `extract_common_adsl_variables()` to streamline the identification and analysis of common ADSL variables across domains. This tool aids in verifying data consistency and understanding variable reuse in clinical datasets.

# See Also

For more details on extracting Define-XML metadata, see `?extract_variable_info_from_define()`.
